VOCC outline

### 2020-04-24 ms/ms.Rmd
edits to intro outline, methods and results

**Question framing notes:**

-   *Groundfish*

    -   *occupy a relatively homogenous thermal environment*

        -   *predict less tendency to move in response to climate*

    -   *many migrate and school*

    -   *high possibility of movement in response to climate*

-   *Stock assessments assume stationary distributions*

-   *Choices of marine protected areas*

-   *Approaches used often consider the whole of species ranges and apply forward projections assuming that species responses to climate variables will remain the same through time.*

@@

-   **Range boundaries** ...

-   **Center of gravity** -- commonly used to assess responses...

@@

-   ... Could complicate interpretation of indices such as COG?

-   **Local climate velocities** are one way of quantifying these changes at finer spatial scales.

@@

*DO I need to talk about trends here, if trend analysis is included?*

@@

@@

@@

Fisheries and Oceans Canada conducts fishery-independent bottom trawl surveys with biennial coverage of the continental shelf and upper slope of most of the BC coast.
These surveys follow a random depth stratified design and use the same bottom trawl fishing gear and fishing protocols (TODO: Sinclair et al. 2003).
Surveys are stratified within four separate regions two of which are surveyed each year since 2005 (with surveys also conducted in some regions in 2003 & 2004): Hecate Strait and Queen Charlotte Sound (odd years) and West Coast Vancouver Island and West Coast Haida Gwaii (even years).
Each region is sampled over the same month long period between late May and early August each year.
Only successful tows of 15--20 min duration are included in analyses and total biomass of each species captured was converted to a biomass density estimate based on the speed, distance covered and net opening.
Catches of between 10--50 individuals are weighed and measured individually, while larger catches are subsampled and for commercially important species data on maturity and age are also collected.

@@

We chose 36 species of Pacific groundfish most regularly encountered on fisheries-independent bottom trawl surveys and relatively widely distributed within Canadian waters.
Because ontogenetic shifts in habitat, particularly depth, are well documented in groundfish species [@mindel2016; @barbeaux2018; @li2019], we estimate biomass densities separately for mature and immature size classes whenever possible (maturity is not assessed for certain Chondrichthians, or when catches are particularly low)
To do this, we first estimate length at 50% maturity (defined by gonadal development stages of immature or maturing vs. mature, ripe, or spent) using ogives fit as sex-specific logistic regressions to individual specimens.
When sufficient samples were available in all years, a random effect of year was included to allow for temporal change in size at maturity (TODO: REF) and these year specific thresholds were used to split observed catches.
For those species lacking maturity data in some years, a random effect of tow id was included instead, but catches were split based on the global estimate.
For all catches exceeding the species-specific threshold count of individual fish, either a random subset of roughly 50 fish, or all individuals will be measured for both length and weight (occasional missing weights were estimated from $\log(W_i) \sim \mathrm{Student-t} (df = 3, \log(a) + b \log(L_i), \sigma)$, with $W_i$ and $L_i$ representing the weight and length for fish $i$ and $\sigma$ representing the observation error scale).
We then calculate the summed weight of all measured which exceeded the length at 50% maturity threshold, divided by the total weight of all measured fish, and multiply this ratio by the estimated biomass density for each tow.
For each tow which resulted in a catch too small for detailed measurements to have been taken, we apply the mean ratio from all measured tows to estimate mature biomass.
Immature density was estimated as the difference between the total density and mature density.

@@

A Seabird Electronics 19plus conductivity, temperature, and depth recorder (CTD) with dissolved oxygen (DO) sensor has also been successfully deployed on most tows since 2008.
Therefore, we modeled bottom temperature and DO levels from CTD sensor data collected on each survey tow using a similar geostatistical model as for biomass.
The approach for these models differed only be being Gaussian and by estimating the spatiotemporal random fields using an AR1 process.
Because DO levels cannot be negative and are known to be influenced be both water temperature and seasonal biological processes (REF), we predict log(DO) with added fixed effects for day of year (centered and scaled) and temperature (quadratic of scaled values similar to depth).
Bottom temperature data (but not DO) was also collected on tows prior to 2008, so this data was also included in the model in order to better inform the static spatial field component of the model.

@@

Across species, the most common patterns were that 1) the impact of warming on biomass change was more negative (or at least less positive) in locations with higher mean temperatures than in cooler locations (e.g. Widow Rockfish, mature Shortspine Thornyhead, and mature English and Flathead Sole), 2) the impact of increasing DO levels was more negative in locations with lower mean DO levels than in locations with higher mean DO levels (e.g. Dogfish, mature Sablefish, Widow Rockfish, and English and Flathead Sole), and 3) overall immature abundances were more likely than mature abundances to show negative relationships (eg. Sablefish, Widow Rockfish, and Dover Sole (Fig. \@ref(fig:trend-chopsticks)). 
<!-- TODO: checking this with model including age as fixed effect (mature biomass as the intercept)... --> 
There were no strong patterns of similarity amoungst species belong to the same genus or family (TODO: should we mention here that they were included as random effects in trial models, but not the final model?) as illustrated by relationships with biotic trends that were negative at high temperatures (Canary Rockfish and English Sole) or positive at low temperatures (eg. Widow Rockfish and Flathead Sole) occuring in members of both the largest families represented in our analysis, Sebastidae (rockfish) and Pleuronectidae (righteye flounders; Fig. Fig. \@ref(fig:worm-plots and \@ref(fig:species-map-chop-panel)). 
Likewise, most species in all families showed relatively little impact of DO change when DO levels were high, but a potentially counter-intuitive decline in biomass if DO increased when mean DO was at it's lowest (\@ref(fig:trend-chopsticks). 
However, several species, mostly rockfish, do show increasing biomass with increasing DO across their distributions (mature Bocaccio, Greenstripe, Redstripe, Yelloweye, Yellowmouth and Yellowtail Rockfishes, and immature Sharpchin Rockfish and Sablefish).
The few exceptions to these pattern include a strong positive effect of increasing DO at low mean levels for adult Splitnose Rockfish, and both maturity classes of Shortspine Thornyheads, all groups occupying some of the deeper habitats (Fig. \@ref(fig:slope-by-depth)). 
However, three species with deeper mean occurence show the more common pattern (mature Sablefish and all of the Rougheye/Blackspotted Rockfish complex), or no relationship at all with DO change (Shortraker Rockfish).

>>>>

-   **Why study species responses to climate (Ecological and management perspectives)?**

    -   Stock assessments assume stationary distributions
    -   Choices of marine protected areas

@@

Approaches used often consider the whole of species ranges and apply forward projections assuming that species responses to climate variables will remain the same through time.

-   ***Range boundaries*** ...
-   ***Center of gravity*** -- commonly used to assess responses...

@@

    -  Could complicate interpretation of indices such as COG?

-   ***Local climate velocities*** are one way of quantifying these changes at finer spatial scales.

@@

<!-- TODO: DO I need to talk about trends here, if trend analysis is included? -->

@@

- 	**Why groundfish in BC are a good case study / application**

    -   occupy a relatively homogenous thermal environment
        -   predict less tendency to move in response to climate
    -   many migrate and school
    -   high possibility of movement in response to climate

- 	**Past work on VOCC in BC groundfish, prevailing hypotheses, limitations/unanswered questions**

@@

<!-- TODO: Consider starting the methods with one short paragraph giving an overview of the methods -->

@@

We analyze abundance, distribution, and morphometric data for 36 species of Pacific groundfish most regularly encountered on fisheries-independent bottom trawl surveys and relatively widely distributed within Canadian waters. 
Surveys are stratified within four separate regions two of which are surveyed each year since 2005 (with surveys also conducted in some regions in 2003 & 2004): Hecate Strait and Queen Charlotte Sound (odd years) and West Coast Vancouver Island and West Coast Haida Gwaii (even years). 
Each region is sampled over the same month long period between late May and early August each year.
Combined these surveys cover most of the upper continental slope and shelf in Canada and share the same random depth stratified design, fishing gear, and fishing protocols (TODO: Sinclair et al. 2003).
Only successful tows of 15--20 min duration are included in analyses and total biomass of each species captured was converted to a biomass density estimate based on the speed, distance covered and net opening.
Catches of between 10--50 individuals are weighed and measured individually, while larger catches are subsampled and for commercially important species data on maturity and age are also collected.

@@

Because ontogenetic shifts in habitat, particularly depth, are well documented in groundfish species [@mindel2016; @barbeaux2018; @li2019], we estimate biomass densities separately for mature and immature size classes whenever possible (maturity is not assessed for certain Chondrichthians, or when catches are particularly low).
To do this, we first estimate length at 50% maturity (defined by gonadal development stages of immature or maturing vs. mature, ripe, or spent) using ogives fit as sex-specific logistic regressions to individual specimens.
When sufficient samples were available in all years, a random effect of year was included to allow for temporal change in size at maturity (TODO: REF) and these year specific thresholds were used to split observed catches.
For those species lacking maturity data in some years, a random effect of tow id was included instead, but catches were split based on the global estimate.
For all catches exceeding the species-specific threshold count of individual fish, either a random subset of roughly 50 fish, or all individuals will be measured for both length and weight (occasional missing weights were estimated from $\log(W_i) \sim \mathrm{Student-t} (df = 3, \log(a) + b \log(L_i), \sigma)$, with $W_i$ and $L_i$ representing the weight and length for fish $i$ and $\sigma$ representing the observation error scale).
We then calculate the summed weight of all measured which exceeded the length at 50% maturity threshold, divided by the total weight of all measured fish, and multiply this ratio by the estimated biomass density for each tow.
For each tow which resulted in a catch too small for detailed measurements to have been taken, we apply the mean ratio from all measured tows to estimate mature biomass.
Immature density was estimated as the difference between the total density and mature density.

@@

We modeled bottom temperature and DO levels from sensor data collected by Seabird Electronics 19plus conductivity, temperature, and depth recorder (CTD) with dissolved oxygen (DO) sensor was successfully deployed on most survey tows since 2008. Our modeling approach was similar to that used for biomass densities except that the distribution was Gaussian and we estimated the spatiotemporal random fields using an AR1 process.
Because DO levels cannot be negative and are known to be influenced be both water temperature and seasonal biological processes (REF), we predicted log(DO) with fixed effects for day of year (centered and scaled) and temperature (quadratic of scaled values similar to depth).
Bottom temperature data (but not DO) was also collected on tows prior to 2008, so this data was also included in the model in order to better inform the static spatial field component of the model.

@@

TODO: Add comment on significant negative global effects of both temperature and DO change, and negative interact for temperature. Maybe also add a simplier version of (Fig. \@ref(fig:supp-global-coefs)) to paper?

Across species, the most common patterns were that 
1) the impact of warming on biomass change was more negative (or at least less positive) in locations with higher mean temperatures than in cooler locations (e.g. Widow Rockfish, mature Shortspine Thornyhead, and mature English and Flathead Sole), 
2) the impact of increasing DO levels was more negative in locations with lower mean DO levels than in locations with higher mean DO levels (e.g. Dogfish, mature Sablefish, Widow Rockfish, and English and Flathead Sole) <!-- Note that negative effect of DO change is sig overall, but not interaction... so may change this? -->, and 
3) overall immature abundances showed a tendency to respond more positively to both mean DO and changes in DO (Fig. \@ref(fig:supp-global-coefs)). 
There were no strong patterns of similarity amoungst species belonging to the same genus or family (TODO: should we mention here that they were included as random effects in trial models, but not the final model?) as illustrated by relationships with biotic trends that were negative at high temperatures (Canary Rockfish and English Sole) or positive at low temperatures (eg. Widow Rockfish and Flathead Sole) occuring in members of both the largest families represented in our analysis, Sebastidae (rockfish) and Pleuronectidae (righteye flounders; Fig. Fig. \@ref(fig:worm-plots and \@ref(fig:species-map-chop-panel)). 
Likewise, most species in all families showed relatively little impact of DO change when DO levels were high, but a potentially counter-intuitive decline in biomass if DO increased when mean DO was at it's lowest (\@ref(fig:trend-chopsticks). 
However, several species, mostly rockfish, do show increasing biomass with increasing DO across their distributions (mature Bocaccio, Greenstripe, Redstripe, Yelloweye, Yellowmouth and Yellowtail Rockfishes, and immature Sharpchin Rockfish and Sablefish).
The few exceptions to these pattern include a strong positive effect of increasing DO at low mean levels for adult Splitnose Rockfish, and both maturity classes of Shortspine Thornyheads, all groups occupying some of the deeper habitats (Fig. \@ref(fig:slope-by-depth)). 
However, three species with deeper mean occurence show the more common pattern (mature Sablefish and all of the Rougheye/Blackspotted Rockfish complex), or no relationship at all with DO change (Shortraker Rockfish).
